---
title: "Behavioral Risk Factor CDC Project"
author: "Scott Povlak"
date: "2023-08-16"
output: html_document
---
Load tidyverse package
```{r}
suppressPackageStartupMessages(library(tidyverse))
install.packages('ggplot2')
```

Read csv file into R
```{r}
brf <- read_csv("BRFSS2015_650.csv")
head(brf)
```

Count of people who think their general health is poor
```{r}
gen_health_poor <- brf %>%
  filter(GENHLTH == 5) %>%
  summarize(brf = n())
gen_health_poor
```

Create distribution for how many days people thought physical health was poor
```{r}
poor_physical <- brf %>%
  filter(PHYSHLTH != 88 & PHYSHLTH != 77 & PHYSHLTH != 99 & !is.na(PHYSHLTH))

ggplot(data = poor_physical, aes(x = PHYSHLTH)) +
  geom_histogram(binwidth = 1, fill = 'blue', color = 'black') + 
  labs(x = "# of Days with Poor Physical Health", y = "Frequency", title = "Distrubtion of Physical Health")
```
Create distribution for how many days people thought mental health was poor
```{r}
poor_mental <- brf %>%
  filter(MENTHLTH != 88 & MENTHLTH != 77 & MENTHLTH != 99)

ggplot(data = poor_mental, aes(x = MENTHLTH)) +
  geom_histogram(binwidth = 1, fill = 'lightblue', color = 'black') + 
  labs(x = "# of Days with Poor Mental Health", y = "Frequency", title = "Distrubtion of Mental Health")
```
Compare mean and standard deviation of physical health comparing those who are
verterans and those who are not
```{r}
vet_physhlth <- brf %>%
  filter(VETERAN3 %in% c(1, 2)) %>%
  mutate(VETERAN3 = case_when(
    VETERAN3 == 1 ~ "Yes",
    VETERAN3 == 2 ~ "No",
  )) %>%
  group_by(VETERAN3) %>%
  mutate(PHYSHLTH = case_when( 
    PHYSHLTH == "88" ~ 0,
    PHYSHLTH == "77" ~ NaN,
    PHYSHLTH == "99" ~ NaN,
    PHYSHLTH == "BLANK" ~ NaN,
    TRUE ~ as.numeric(PHYSHLTH)  
    )) %>%
  summarize(mean_physical = round(mean(PHYSHLTH, na.rm = TRUE), 2),
            sd_physical = round(sd(PHYSHLTH, na.rm = TRUE), 2)) %>%
  as.data.frame()
  vet_physhlth
```
Summary statistics for age told that a person was diabetic
```{r}
diab_filter <- brf %>%
  filter(DIABAGE2 != 98 & DIABAGE2 != 99 & DIABAGE2 != "BLANK")
summary_stats = diab_filter %>%
  summarise(
    mean_age_diab = mean(DIABAGE2),
    med_age_diab = median(DIABAGE2),
    sd_age_diab = sd(DIABAGE2)
  )
```
Prediction of number of days that mental health was not good from employment status
```{r}
filter_employ <- brf %>%
  filter(EMPLOY1 != 9, !(MENTHLTH %in% c(99, 77)))

filter_employ2 <- filter_employ %>%
  mutate(EMPLOY1 = factor(EMPLOY1),
         MENTHLTH = ifelse(MENTHLTH == 88, 0, MENTHLTH))

lm_model <- lm(MENTHLTH ~ EMPLOY1, data = filter_employ2)
menhlth_employ <- summary(lm_model)
```
ANOVA comparing how many times a week they participated in physical activity against 
employment status
```{r}
phy_act_employ <- brf %>%
  mutate(EMPLOY1 = factor(EMPLOY1)) %>%
  filter(
    EMPLOY1 != 9,
    !(EXEROFT1 %in% c(777, 999)),
    EXEROFT1 != "BLANK",
    !is.na(EXEROFT1),
    !(EXEROFT1 %in% seq(201, 299)) 
  ) %>%
  aov(EXEROFT1 ~ EMPLOY1, data = .)

exeroft_employ <- summary(phy_act_employ)
```
Box plot for ANOVA above
```{r}
ggplot(data = phy_act_employ, aes(x = EMPLOY1, y = as.numeric(EXEROFT1))) +
  geom_boxplot(fill = "blue", alpha = 0.7) +
  labs(x = "Employment Status", y = "Physical Activity", title = "ANOVA: Box Plot of Physical Activity by Employment Status")
```
